---
title: "Project-Code"
author: "Dennis Goldenberg"
date: "2024-11-27"
output: pdf_document
---
# Predicting CAT Bond Expected Market Yield using Time Series Data
```{r, include = TRUE}
library(readxl)
suppressWarnings(library(quantmod))
suppressWarnings(library(ggplot2))
```

## Reading in Data
```{r, include = TRUE}
#Response variable
yield_w_EL <- read_xlsx(path = "data/yield_with_EL.xlsx",
                        skip = 2)
yield_w_EL$Date <- as.Date(yield_w_EL$Date)

#Credit Spread (for potential modeling of Reinsurance premium)
credit_spread <- read_xls(path = "data/Credit_Spread_BofA.xls",
                          skip = 10)
credit_spread$observation_date <- as.Date(credit_spread$observation_date)
colnames(credit_spread) <- c("Date", "Spread")

#S & P Information
S_and_P <- read_xls(path = "data/SP500.xls",
                    skip = 10)
S_and_P$observation_date <- as.Date(S_and_P$observation_date)
colnames(S_and_P) <- c("Date", "SP500")

#EAFE Index
EAFE <- read.csv("data/EAFE International Index value.csv")
```


```{r, include = TRUE}
ggplot(data = yield_w_EL) + geom_line(aes(x = Date,
                                          y = `Money Market rate (USD)`))
```
```{r, include = TRUE}
ggplot(data = yield_w_EL) + geom_line(aes(x = Date,
                          y = `weekly average reinsurance market yield (USD)`))
```
```{r, include = TRUE}
ggplot(data = credit_spread) + geom_line(aes(x = Date,
                                       y = Spread/100))
```

```{r, include = TRUE}
ggplot() + geom_line(data = yield_w_EL,
                      mapping = aes(x = Date,
                      y = `weekly average reinsurance market yield (USD)`,
                      color = 'Reinsurance Market Yield')) + 
  geom_line(data = credit_spread, aes(x = Date, y = Spread/100, 
                                      color = 'Credit Spread')) +
  labs(colour = 'Quantity') + ylab('Value') + 
  ggtitle("Credit spread & Expected Reinsurance Market Yield over time") + 
  theme(legend.position = "inside", legend.position.inside =  c(0.15, .85),
        plot.title = element_text(hjust = .5))
```

```{r, include = TRUE}
ggplot() + geom_line(data = yield_w_EL,
                      mapping = aes(x = Date,
                      y = `weekly average reinsurance market yield (USD)`,
                      color = 'Reinsurance Market Yield')) + 
  geom_line(data = yield_w_EL, aes(x = Date, y = `Expected Loss (EL)`, 
                                      color = 'Expected Loss')) +
  labs(colour = 'Quantity') + ylab('Value') + 
  ggtitle("Credit spread & Expected Reinsurance Market Yield over time") + 
  theme(legend.position = "inside", legend.position.inside =  c(0.15, .85),
        plot.title = element_text(hjust = .5))
```

## ACF/Stationarity Analysis

### Response variable (reinsurance market yield)
```{r, include = TRUE}
acf(yield_w_EL$`weekly average reinsurance market yield (USD)`,
    lag.max = 50)
```
Clearly, our response variable is a random walk.

```{r, include = TRUE}
acf(diff(yield_w_EL$`weekly average reinsurance market yield (USD)`))
```

```{r, include = TRUE}
pacf(diff(yield_w_EL$`weekly average reinsurance market yield (USD)`))
```

Looks to be an AR(1) process.

```{r, include = TRUE}
ggplot() + geom_line(aes(x = yield_w_EL$Date[2:length(yield_w_EL$Date)],
                         y = diff(yield_w_EL$`weekly average reinsurance market yield (USD)`)))
```

### BoFA Credit Spread
```{r, include = TRUE}
acf(credit_spread$Spread, lag.max = 50)
```
Also, clearly a random walk:
```{r, include = TRUE}
acf(diff(credit_spread$Spread/100))
```
```{r, include = TRUE}
pacf(diff(credit_spread$Spread/100))
```
```{r, include = TRUE}
ggplot() + geom_line(aes(x = credit_spread$Date[2:length(credit_spread$Date)],
                         y = diff(credit_spread$Spread/100)))
```

```{r, include = TRUE}
acf(yield_w_EL$`Expected Loss (EL)`, lag.max = 100)
```
Also, clearly a random walk.
```{r, include = TRUE}
acf(diff(yield_w_EL$`Expected Loss (EL)`))
```
Little evidence that serial correlation remains.


## Relationships between variables analysis
```{r, include=TRUE}
ccf(diff(yield_w_EL$`weekly average reinsurance market yield (USD)`),
    diff(credit_spread$Spread/100),
    lag.max = 200)
```



# Analyzing Correlation with S&P and EAFE
```{r, include = TRUE}
ggplot(data = S_and_P) + geom_line(aes(x = Date, y = SP500))
```

```{r, include = TRUE}
q1_2015_ret <- (2074.828 - 2072.005)/2072.005
q1_2015_ret
```

```{r, include = TRUE}
yield_w_EL[c("Date", "Total Coupon CAT Bond Market (USD)")]
```

```{r, include = TRUE}
(0.04529922 - 0.03767351)/0.03767351
```


