---
title: "Project-Code"
author: "Dennis Goldenberg"
date: "2024-11-27"
output: pdf_document
---
# Predicting CAT Bond Expected Market Yield using Time Series Data
```{r, include = TRUE}
library(readxl)
suppressWarnings(library(quantmod))
suppressWarnings(library(ggplot2))
library(stats)
```

## Reading in Data
```{r, include = TRUE}
#Response variable
yield_w_EL <- read_xlsx(path = "data/yield_with_EL.xlsx",
                        skip = 2)
yield_w_EL$Date <- as.Date(yield_w_EL$Date)

#Credit Spread (for potential modeling of Reinsurance premium)
credit_spread <- read_xls(path = "data/Credit_Spread_BofA_weekly.xls",
                          skip = 10)
credit_spread$observation_date <- as.Date(credit_spread$observation_date)
colnames(credit_spread) <- c("Date", "Spread")

#Break-Even Interest Rate (expected measure of inflation)
BEIR <- read_xls(path = "data/BEIR.xls",
                          skip = 10)
BEIR$observation_date <- as.Date(BEIR$observation_date)
colnames(BEIR) <- c("Date", "BEIR")

#S & P Information
S_and_P <- read_xls(path = "data/SP500.xls",
                    skip = 10)
S_and_P$observation_date <- as.Date(S_and_P$observation_date)
colnames(S_and_P) <- c("Date", "SP500")

#EAFE Index
EAFE <- read.csv("data/EAFE International Index value.csv")
```


```{r, include = TRUE}
ggplot(data = yield_w_EL) + geom_line(aes(x = Date,
                                          y = `Money Market rate (USD)`))
```
```{r, include = TRUE}
ggplot(data = yield_w_EL) + geom_line(aes(x = Date,
                          y = `weekly average reinsurance market yield (USD)`))
```
```{r, include = TRUE}
ggplot(data = credit_spread) + geom_line(aes(x = Date,
                                       y = Spread/100))
```

```{r, include = TRUE}
ggplot() + geom_line(data = yield_w_EL,
                      mapping = aes(x = Date,
                      y = `weekly average reinsurance market yield (USD)`,
                      color = 'Reinsurance Market Yield')) + 
  geom_line(data = credit_spread, aes(x = Date, y = Spread/100, 
                                      color = 'Credit Spread')) +
  labs(colour = 'Quantity') + ylab('Value') + 
  scale_x_continuous("Date", breaks = c(14975, 15340, 15705,
                                        16070,16435, 16800,
                                        17165, 17530, 17895,
                                        18260, 18625, 18990,
                                        19355, 19720),
                     labels = c(2011, 2012, 2013, 2014, 2015, 2016,
                                2017, 2018, 2019, 2020, 2021, 2022,
                                2023, 2024)) + 
  ggtitle("Credit spread & Expected Reinsurance Market Yield over time") + 
  theme(legend.position = "inside", legend.position.inside =  c(0.15, .85),
        plot.title = element_text(hjust = .5))
```
```{r, include = TRUE}
ggplot() + geom_line(data = yield_w_EL,
                      mapping = aes(x = Date,
                      y = `weekly average reinsurance market yield (USD)`,
                      color = 'Reinsurance Market Yield')) + 
  geom_line(data = BEIR, aes(x = Date, y = BEIR/100, 
                                      color = 'BEIR')) +
  labs(colour = 'Quantity') + ylab('Value') + 
  scale_x_continuous("Date", breaks = c(14975, 15340, 15705,
                                        16070,16435, 16800,
                                        17165, 17530, 17895,
                                        18260, 18625, 18990,
                                        19355, 19720),
                     labels = c(2011, 2012, 2013, 2014, 2015, 2016,
                                2017, 2018, 2019, 2020, 2021, 2022,
                                2023, 2024)) + 
  ggtitle("Credit spread & Expected Reinsurance Market Yield over time") + 
  theme(legend.position = "inside", legend.position.inside =  c(0.15, .85),
        plot.title = element_text(hjust = .5))
```




```{r, include = TRUE}
ggplot() + geom_line(data = yield_w_EL,
                      mapping = aes(x = Date,
                      y = `weekly average reinsurance market yield (USD)`,
                      color = 'Reinsurance Market Yield')) + 
  geom_line(data = yield_w_EL, aes(x = Date, y = `Expected Loss (EL)`, 
                                      color = 'Expected Loss')) +
  labs(colour = 'Quantity') + ylab('Value') + 
   scale_x_continuous("Date", breaks = c(14975, 15340, 15705,
                                        16070,16435, 16800,
                                        17165, 17530, 17895,
                                        18260, 18625, 18990,
                                        19355, 19720),
                     labels = c(2011, 2012, 2013, 2014, 2015, 2016,
                                2017, 2018, 2019, 2020, 2021, 2022,
                                2023, 2024)) + 
  ggtitle("Credit spread & Expected Reinsurance Market Yield over time") + 
  theme(legend.position = "inside", legend.position.inside =  c(0.15, .85),
        plot.title = element_text(hjust = .5))
```

## ACF/Stationarity Analysis

### train/test split
```{r, include = TRUE}
credit_spread_daily <- read_xls(path = "data/Credit_Spread_BofA_daily.xls",
                          skip = 10)
credit_spread_daily$observation_date <- as.Date(credit_spread_daily$observation_date)
colnames(credit_spread_daily) <- c("Date", "Spread")
full_data <- merge(yield_w_EL, credit_spread_daily, by = 'Date')
full_data['Spread'] <- round(full_data['Spread']/100,6)
train_data <- full_data[c(which(full_data$Date <= as.Date("2020-10-08"))),]
test_data <- full_data[c(which(full_data$Date > as.Date("2020-10-08"))),]
good_col_names <- c("Date", "Yield", "MMR", "EL", "Total_Yield","Spread")

colnames(train_data) <- good_col_names
colnames(test_data) <- good_col_names
```

```{r, include = TRUE}
full_data
```



### Response variable (reinsurance market yield)
```{r, include = TRUE}
acf(train_data$Yield,
    lag.max = 100)
```
Clearly, our response variable is a random walk.

```{r, include = TRUE}
acf(diff(train_data$Yield))
```

```{r, include = TRUE}
pacf(diff(train_data$Yield))
```
```{r, include = TRUE}
lag_1 <- pacf(diff(train_data$Yield),
     plot = F)[1]
lag_1
```


Looks to be an AR(1) process. Specifically, the formula looks to be:
\begin{equation*}
\text{yield}_t = -.367 * \text{yield}_{t - 1} + w_t
\end{equation*}
This can be rewritten as $\theta_1(\mathbf{B})y_t = w_t$, where $\theta_1(\mathbf{B}) = 1 + .367\mathbf{B}$. The root of this equation is $\frac{-1}{.367}$ and $\bigg|\frac{-1}{.367}\bigg| > 1$, so the series is stationary.

```{r, include = TRUE}
ggplot() + geom_line(aes(x = train_data$Date[2:length(train_data$Date)],
                         y = diff(train_data$Yield)))
```

### BoFA Credit Spread
```{r, include = TRUE}
acf(train_data$Spread, lag.max = 50)
```
Also, clearly a random walk:
```{r, include = TRUE}
acf(diff(train_data$Spread))
```
```{r, include = TRUE}
pacf(diff(train_data$Spread/100))
```
Statistically significant lags up to lag 4. I run an ar model, with order.max at 5, to check for stationary:

```{r, include = TRUE}
spread_ar <- ar(diff(train_data$Spread), aic = TRUE, order.max = 5,
                demean = TRUE)
spread_ar
```

```{r, include = TRUE}
polyroot(c(1,spread_ar$ar * -1))
```
All roots are above 1 in magnitude. Therefore, I conclude that the differenced credit spread is stationary.


```{r, include = TRUE}
ggplot() + geom_line(aes(x = train_data$Date[2:length(train_data$Date)],
                         y = diff(train_data$Spread)))
```

```{r, include = TRUE}
acf(train_data$EL, lag.max = 100)
```
Also, clearly a random walk.
```{r, include = TRUE}
acf(diff(train_data$EL))
```

```{r, include = TRUE}
pacf(diff(train_data$EL))
```
```{r, include = TRUE}
EL_ar <- ar(diff(train_data$EL), aic = TRUE, order.max = 10,
                demean = TRUE)
EL_ar
```
```{r, include = TRUE}
polyroot(c(1, EL_ar$ar * -1))
```
All roots have magnitude greater than 1. Therefore, we can conclude that this series is stationary as well.



## Relationships between variables analysis

### Credit Spread and Yield
```{r, include=TRUE}
ccf(diff(train_data$Spread),
    diff(train_data$Yield),
    lag.max = 100)
```
```{r, include = TRUE}
ccf(diff(train_data$Spread),
    diff(train_data$Yield),
    lag.max = 200, plot = FALSE)[seq(-40, -30, 1)]
```
Strong cross-correlation at about 36 weeks (about 9 months) back.


### EL and Yield
```{r, include = TRUE}
ccf(diff(train_data$EL),
    diff(train_data$Yield),
    lag.max = 100)

```
```{r, include = TRUE}
cor(train_data[,c("Yield",
                  "Spread", "EL")])
```
```{r, include = TRUE}
ggplot(data = train_data) + geom_point(aes(x = EL, y = Yield))
```



Strong correlation, EL should be an exogenous variable.

## Generating the ARIMA model

### Model 1: ARIMA(1,1,0)
```{r, include = TRUE}
m1 <- arima(train_data$Yield, order = c(1,1,0), include.mean = TRUE)
pred_m1 <- predict(m1, n.ahead = length(test_data$Yield))
ggplot() + geom_line(aes(x = test_data$Date, y = test_data$Yield, 
                         color = 'Actual')) + 
  labs(colour = 'Quantity') + xlab('Date') + ylab('Yield') +
  geom_line(aes(x = test_data$Date, y = pred_m1$pred,
                color = 'predicted'))
```
```{r, include = TRUE}
m1
```

### Model 2: ARIMA(1,1,0) w/ EL
```{r, include = TRUE}
m2 <- arima(train_data$Yield, order = c(1,1,0), 
            xreg = train_data$EL, 
            include.mean = TRUE)
m2
```
```{r, include = TRUE}
pred_m2 <- predict(m2, n.ahead = length(test_data$Yield),
                   newxreg = test_data$EL)
ggplot() + geom_line(aes(x = test_data$Date, y = test_data$Yield, 
                         color = 'Actual')) + 
  labs(colour = 'Quantity') + xlab('Date') + ylab('Yield') +
  geom_line(aes(x = test_data$Date, y = pred_m2$pred,
                color = 'predicted'))
```
### Model 3: ARIMA(1,1,0) w/ EL and Credit Spread

```{r, include = TRUE}
spread_lag_36 <- full_data$Spread[seq(1, length(train_data$Spread) - 36)]
spread_lag_36_test <- full_data$Spread[seq(length(train_data$Spread) - 36,
                                        length(diff(full_data$Spread)) - 36)]
yield_for_m3 <- train_data$Yield[seq(37,
                                length(train_data$Yield))]
EL_for_m3 <-  train_data$EL[seq(37, length(train_data$EL))]
train_data_m3 <- as.data.frame(alist(
  Yield = yield_for_m3, EL = EL_for_m3, Spread = spread_lag_36))
m3 <- arima(train_data_m3$Yield, order = c(1,1,0),
            xreg = as.matrix.data.frame(train_data_m3[,c("EL","Spread")]))
m3
```

```{r, include = TRUE}
test_data_m3 <- as.data.frame(alist(
  Yield = test_data$Yield,
  EL = test_data$EL,
  Spread_lag_36 = spread_lag_36_test
))
```

```{r, include = TRUE}
pred_m3 <- predict(m3, n.ahead = length(test_data$Yield),
                   newxreg = as.matrix.data.frame(test_data_m3[,c("EL","Spread_lag_36")]))
ggplot() + geom_line(aes(x = test_data$Date, y = test_data$Yield, 
                         color = 'Actual')) + 
  geom_line(aes(x = test_data$Date, y = pred_m3$pred,
                color = 'predicted')) + 
  labs(colour = 'Quantity') + xlab('Date') + ylab('Yield')
```

```{r, include = TRUE}
train_pred <- train_data$Yield + m1$residuals
ggplot() + geom_line(aes(x = train_data$Date, y = train_data$Yield)) + 
  geom_line(aes(x = train_data$Date, y = train_pred))
```

# Analyzing Correlation with S&P
```{r, include = TRUE}
percent_changes_sp <- numeric(length(S_and_P$SP500) - 12)
percent_changes_cat <- percent_changes_sp

for (i in 1:(length(S_and_P$SP500) - 12)) {
  percent_changes_sp[i] <- ((S_and_P$SP500[i + 12] - S_and_P$SP500[i]) / S_and_P$SP500[i]) * 100
  percent_changes_cat[i] <- ((yield_w_EL$`Total Coupon CAT Bond Market (USD)`[i + 222 + 12] - yield_w_EL$`Total Coupon CAT Bond Market (USD)`[i + 222]) / yield_w_EL$`Total Coupon CAT Bond Market (USD)`[i + 222]) * 100
}

change_in_yield <- data.frame(
  Date_sp = S_and_P$Date[1:(length(S_and_P$Date) - 12)],
  Date_cat = yield_w_EL$Date[223:(length(S_and_P$Date) - 12 + 223 - 1)],
  Percent_Change_sp = percent_changes_sp,
  Percent_Change_cat = percent_changes_cat
)

head(change_in_yield)
```

```{r, include = TRUE}
ggplot(data = S_and_P) + geom_line(aes(x = Date, y = SP500))
```

```{r, include = TRUE}
q1_2015_ret <- (2074.828 - 2072.005)/2072.005
q1_2015_ret
```

```{r, include = TRUE}
yield_w_EL[c("Date", "Total Coupon CAT Bond Market (USD)")]
```

```{r, include = TRUE}
(0.04529922 - 0.03767351)/0.03767351
```


